---
title: "RWorksheet_Arcena#3a"
author: "John Pierre R. Arceña"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Using Vectors

# There is a built-in vector LETTERS contains the 
# uppercase letters of the alphabet and letters which 
# contains the lowercase letters of the alphabet.

## Based on the above vector LETTERS:

##1a. You need to produce a vector that contains the 
# first 11 letters.
```{r}
LETTERS[1:11]
```

##1b. Produce a vector that contains the odd numbered 
# letters.
```{r}
LETTERS[seq(1, 26, by = 2)]
```

##1c. Produce a vector that contains the vowels
```{r}
LETTERS[LETTERS %in% c("A", "E", "I", "O", "U")]
```

##Based on the above vector letters:

##1d. Produce a vector that contains the last 5 
# lowercase letters.
```{r}
tail(letters, 5)
```

##1e. Produce a vector that contains letters between 
# 15 to 24 letters in lowercase.
```{r}
letters[15:24]
```

## 2. Temperature Data

# Create a vector (not a dataframe) with the average 
# temperatures in April for Tuguegarao City, Manila, 
# Iloilo City, Tacloban, Samal Island, and Davao City.

##2a. What is the R code and its result for creating 
# a character vector for the city/town? Name the object 
# as city.
```{r}
city <- c("Tuguegarao City", "Manila", "Iloilo City", 
          "Tacloban", "Samal Island", "Davao City")
city
```

##2b. The average temperatures in Celcius are 42, 39, 
# 34, 34, 30, and 27 degrees. Name the object as temp.
```{r}
temp <- c(42, 39, 34, 34, 30, 27)
temp
```

##2c. Create a dataframe to combine the city and the 
# temp by using data.frame().
```{r}
city_temp <- data.frame(city, temp)
city_temp
```

##2d. Associate the dataframe by naming the columns 
# using the names() function as City and Temperature.
```{r}
names(city_temp) <- c("City", "Temperature")
city_temp
```

##2e. Print the structure by using str() function. 
# Describe the output.
```{r}
str(city_temp)
```

##2f. From the answer in d, what is the content of 
# row 3 and row 4?
```{r}
city_temp[3:4, ]
```

##2g. Display the city with highest temperature and 
# the city with the lowest temperature.
```{r}
# City with highest temperature
city_temp[city_temp$Temperature == max(city_temp$Temperature), ]

# City with lowest temperature
city_temp[city_temp$Temperature == min(city_temp$Temperature), ]
```

## Using Matrices

##2. Create a matrix of one to eight and eleven to 
# fourteen with four columns and three rows.

##2a. What will be the R code for the #2 question 
# and its result?
```{r}
m <- matrix(c(1:8, 11:14), nrow = 3, ncol = 4)
m
```

##2b. Multiply the matrix by two. What is its R code 
# and its result?
```{r}
m * 2
```

##2c. What is the content of row 2? What is its R code?
```{r}
m[2, ]
```

##2d. What will be the R code if you want to display 
# the column 3 and column 4 in row 1 and row 2?
```{r}
m[1:2, 3:4]
```

##2e. What is the R code if you want to display only 
# the columns in 2 and 3, row 3?
```{r}
m[3, 2:3]
```

##2f. What is the R code if you want to display only 
# the columns 4?
```{r}
m[, 4]
```

##2g. Name the rows as isa, dalawa, tatlo and columns 
# as uno, dos, tres, quatro.
```{r}
rownames(m) <- c("isa", "dalawa", "tatlo")
colnames(m) <- c("uno", "dos", "tres", "quatro")
m
```

##2h. Reshape the matrix by assigning a new dimension 
# with dim(). New dimensions should have 2 columns and 
# 6 rows.
```{r}
dim(m) <- c(6, 2)
m
```

## Using Arrays

##3. An array contains 1, 2, 3, 6, 7, 8, 9, 0, 3, 4, 5, 1

##3a. Create an array for the above numeric values. 
# Each value will be repeated twice in a three-dimensional 
# array with 4 columns and 2 rows.
```{r}
values <- c(1, 2, 3, 6, 7, 8, 9, 0, 3, 4, 5, 1)
arr <- array(rep(values, 2), dim = c(2, 4, 3))
arr
```

##3b. How many dimensions do your array have?
```{r}
dim(arr)
```

# #3c. Name the rows as lowercase letters and columns 
# as uppercase letters starting from A. The array names 
# should be 1st-Dimensional Array, 2nd-Dimensional Array, 
# and 3rd-Dimensional Array.
```{r}
dimnames(arr) <- list(
  c("a", "b"),
  c("A", "B", "C", "D"),
  c("1st-Dimensional Array", "2nd-Dimensional Array", 
    "3rd-Dimensional Array")
)
arr
```